---
title: "Final Project"
author: "Hannah Wilder and Chathura Gunasekara"
date: "April 9, 2016"
output: pdf_document
geometry: margin=.5in
---


Change working directory
```{r setup, include=FALSE}
library(TSA)

#Hannah
setwd( "C:/Users/Hannah/Downloads/2016SpringTextbooks/TimeSeriesAnalysis/FinalProject/TimeSeriesFinalProject")

#Chathura
#test
#setwd("./")
```


#Load data (assumes file is in working directory)
```{r}
#load the data
beerData<-read.csv("monthly-beer-production-in-austr.csv")

#cut off the last row which is NA
beerData<-beerData[-nrow(beerData),]
colnames(beerData)<-c("Month", "Production")

#Get a five number summary of the data
summary(beerData$Production)

#turn into time series also hold back the last two years of data for forecasting
beerTS<-ts(beerData[1:(nrow(beerData)-24),2], frequency=12, start=c(1956,1))
beer_forecast<-ts(beerData[(nrow(beerData)-23):nrow(beerData), 2], start=c(1993,9), frequency=12)

adf.test(beerTS)
pp.test(beerTS)
```

#Plot population data
```{r}
library(TSA)
par(mfrow=c(1,1))
plot(beerTS, main="Beer Production in Australia by Month")

plot(beerTS, main="Beer Production in Australia by Month (seasons marked)", type="l")
points(y=beerTS, x=time(beerTS), pch=as.vector(season(beerTS)))

require(fpp)
seasonplot(beerTS,year.labels=TRUE,ylab="megaliters",main="Seasonal plot: quarterly beer production", col=rainbow(20), pch=19)

plot(log(beerTS), main="Logged Beer Production in Australia by Month")

plot(beer_forecast, main="Beer production values to be forecasted")

logBeer<-log(beerTS)
```

#Investigate possible relationship with population data
```{r, fig.width=4, fig.height=2}
#load population data
library(reshape)
library(ggplot2)

#Clean up population data
pop_totalData<-t(read.csv("Pop_total.csv", row.names=1))
dropCols<-colnames(pop_totalData) %in% c("Unspecified","Period not indicated")
rownames(pop_totalData)<-c(1921:2011)
pop_totalDataLong<-pop_totalData[,!dropCols]
pop_totalData<-pop_totalData[paste(1956:1995),!dropCols]

#Aggregate beer data
beerYear<-seq(from=1956, to=1996, by=1)
beerYear<-rep(beerYear, each=12)
beerYear<-beerYear[1:nrow(beerData)]
beerAg<-aggregate(beerData[,2], FUN=mean, by=list(year=beerYear))

#Attach to beer data
beerPop<-data.frame(cbind(beer=beerAg[,2],pop_totalData))
beerPopScale<-scale(beerPop)

beerPopRes<-melt(beerPopScale, variable.name="series")
colnames(beerPopRes)<-c("time", "series", "stddev")

allNames<-colnames(beerPop)[2:length(colnames(beerPop))]

#Plot data for each age group and beer data on same plot
par(mfrow=c(2,2))
for (name in allNames) {
  subset_data<-subset(beerPopRes, beerPopRes$series%in%c("beer", name))
  newPlot<-ggplot(subset_data, aes(time,stddev)) + geom_line(aes(colour = series)) +ggtitle(paste("Pop (age group ",name,") and Beer Prod"))
  print(newPlot)
}
par(mfrow=c(1,1))

#Make a model based on the 15-19 age group
yearModel1<-lm(beer ~ X15.19, data=beerPop)
summary(yearModel1)
plot(ts(residuals(yearModel1), frequency=1, start=c(1956)), main="Residuals from modeling beer production with 15-19", ylab="Residual")
```


#Explore lagged x10.14 data
```{r}

laggedData<-data.frame(beer=beerAg[,2])
models<-list()
modelRsq<-c()
for (lag in 0:8) {
  newColNames<-c(colnames(laggedData), paste("lag", lag, sep=""))
  newLag<-pop_totalDataLong[paste(1956:1995-lag), "10-14"]
  laggedData<-data.frame(laggedData, newLag)
  newModel<-lm(beer ~ newLag, data=laggedData)
  models[[paste("lag", lag, sep="")]]<-newModel
  modelRsq<-c(modelRsq, summary(newModel)$r.squared)
  colnames(laggedData)<-newColNames
  
}

plot(modelRsq, main="R Squared Values for lags of 10-14 age group", xlab="lag number", ylab="R squared value")
abline(v=6, lty=2)
max(modelRsq)

lagDataScale<-scale(laggedData)[,c(1,8)]
lagDataMelt<-melt(lagDataScale, variable.name="series")
colnames(lagDataMelt)<-c("time", "series", "stddev")

newPlot<-ggplot(lagDataMelt, aes(time,stddev)) + geom_line(aes(colour = series)) +ggtitle(paste("Lags of 10-14 Pop and Beer Prod"))
print(newPlot)

plot(ts(residuals(models[["lag6"]]), frequency=1, start=c(1956)), main="Residuals from modeling beer production with 10-14 lag 6", ylab="Residual")
abline(h=0)

```

#Interpolate Monthly Numbers
```{r}
library(zoo)

#Create a vector with missing values for zoo to interpolate
withNA<-c()
for (year in 1:nrow(laggedData)) {
  withNA<-c(withNA, laggedData$lag6[year], rep(NA, 11))
}

#Create a vector with missing values for zoo to interpolate
withNALong<-c()
for (year in 1:nrow(pop_totalDataLong)) {
  withNALong<-c(withNALong, pop_totalDataLong[year, "10-14"], rep(NA, 11))
}


#Interpolate values using zoo library
zooSeries<-zoo(withNA, frequency=12)
wAppx<-na.approx(zooSeries, na.rm=FALSE)
monthlyLag6<-wAppx

#Interpolate long series for forecasting
zooSeriesLong<-zoo(withNALong, frequency=12)
wAppxLong<-data.frame(na.approx(zooSeriesLong, na.rm=FALSE))
rownames(wAppxLong)<-round(seq(from=(1921+6), length.out=nrow(wAppxLong), by=1/12),2)
monthlyLag6Long<-wAppxLong

#Reattach to beer numbers for plotting
beerPopMonth<-data.frame(beer=beerTS, lag6_10_14=monthlyLag6[1:length(beerTS)])
rownames(beerPopMonth)<-round(seq(from=1956, length.out=length(beerTS), by=1/12),2)
scaleMonth<-scale(beerPopMonth)
scaleMonthMelt<-melt(scaleMonth, variable.name="series")
colnames(scaleMonthMelt)<-c("time", "series", "stddev")

#Make a pretty plot
newPlot<-ggplot(scaleMonthMelt, aes(time,stddev)) + geom_line(aes(colour = series)) +ggtitle(paste("Lag 6 of 10-14 Pop and Beer Prod"))
print(newPlot)

#Try a model
monthlyPopModel<-lm(beer ~ lag6_10_14, data=beerPopMonth)
summary(monthlyPopModel)

residFromPop<-ts(residuals(monthlyPopModel), frequency=12, start=c(1956,1))
plot(residFromPop, type="l", main="Plot of residuals from Pop model")
points(y=residFromPop, x=time(residFromPop), pch=as.vector(season(residFromPop)))
abline(h=0)
```

#Define helper functions
```{r}
getModelString<-function(model) {
  modSpec<-model$arma
  modelString<-paste("SARIMA(", modSpec[1],",", modSpec[6], ",", modSpec[2],")(", modSpec[3], ",", modSpec[7], ",", modSpec[4],")[", modSpec[5],"]", sep="")
  
  return(modelString)
}

plotResid<-function(model) {
  residuals<-ts(residuals(model), frequency=12, start=c(1956,1))
  modelString<-getModelString(model)
  par(mfrow=c(1,1))
  acf(residuals, main=paste("ACF of", modelString), lag.max=40, cex=.5)
  pacf(residuals, main=paste("PACF of", modelString), lag.max=40, cex=.5)
  par(mfrow=c(1,1))
}

```

#Decompsing the time series to see trends and patterns
```{r}
decompbeer = decompose (beerTS, type="additive")
plot (decompbeer)

#monthplot(beerTS, main="Decomposition of Series by Month")
```

#Try to figure out deterministic trend
```{r}
t<-1:length(beerTS)
t2<-t^2
t3<-t^3
t4<-t^4
t5<-t^5

quadFit<-lm(beerTS~t+t2)
summary(quadFit)

#### plot the data and the fitted quadratic trend function
plot(x=1:length(beerTS),y=beerTS,type='o',ylab="",xlab="Time - Number of Months Since Jan 1956",main="Quadratic Fit on Beer Production Data")
curve(expr = coef(quadFit)[1]+coef(quadFit)[2]*x+coef(quadFit)[3]*x^2+coef(quadFit)[4]*x^3,lty=1,add = TRUE, col="red")

quadFit_resid<-ts(residuals(quadFit),frequency=12, start=c(1956,1)) 
plot(quadFit_resid, main="Residuals from a Quadratic Trend Fit")
abline(h=0)

cubicFit<-lm(beerTS~t+t2+t3)
summary(cubicFit)

#### plot the data and the fitted quadratic trend function
plot(x=1:length(beerTS),y=beerTS,type='o',ylab="",xlab="Time - Number of Months Since Jan 1956",main="Cubic Fit on Beer Production Data")
curve(expr = coef(cubicFit)[1]+coef(cubicFit)[2]*x+coef(cubicFit)[3]*x^2+coef(cubicFit)[4]*x^3,lty=1,add = TRUE, col="red")

cubicFit_resid<-ts(residuals(cubicFit),frequency=12, start=c(1956,1)) 
plot(cubicFit_resid, main="Residuals from a Cubic Trend Fit")
abline(h=0)

order4polyFit<-lm(beerTS~t+t2+t3+t4)
summary(order4polyFit)

#### plot the data and the fitted 4th order polynomial trend function
plot(x=1:length(beerTS),y=beerTS,type='o',ylab="",xlab="Time - Number of Months Since Jan 1956",main="order4poly Fit on Beer Production Data")
curve(expr = coef(order4polyFit)[1]+coef(order4polyFit)[2]*x+coef(order4polyFit)[3]*x^2+coef(order4polyFit)[4]*x^3+coef(order4polyFit)[5]*x^4,lty=1,add = TRUE, col="red")

order4polyFit_resid<-ts(residuals(order4polyFit),frequency=12, start=c(1956,1)) 
plot(order4polyFit_resid, main="Residuals from a order4poly Trend Fit")
abline(h=0)

order5polyFit<-lm(beerTS~t+t2+t3+t4+t5)
summary(order5polyFit)

#### plot the data and the fitted 5th order polynomial trend function
plot(x=1:length(beerTS),y=beerTS,type='o',ylab="",xlab="Time - Number of Months Since Jan 1956",main="order5poly Fit on Beer Production Data")
curve(expr = coef(order5polyFit)[1]+coef(order5polyFit)[2]*x+coef(order5polyFit)[3]*x^2+coef(order5polyFit)[4]*x^3+coef(order5polyFit)[5]*x^4+coef(order5polyFit)[6]*x^5,lty=1,add = TRUE, col="red")

order5polyFit_resid<-ts(residuals(order5polyFit),frequency=12, start=c(1956,1)) 
plot(order5polyFit_resid, main="Residuals from a order5poly Trend Fit")
abline(h=0)

```


#Fit deterministic cosine trend model
```{r}
startDate<-round(start(beer_forecast)[1]+start(beer_forecast)[2]/12,2)
endDate<-round(end(beer_forecast)[1]+end(beer_forecast)[2]/12,2)
numToFor<-length(beer_forecast)

allBeerData<-ts(c(beerTS, beer_forecast), start=c(1956, 1), frequency=12)

tnew<-1:length(allBeerData)
t2new<-tnew^2
t3new<-tnew^3
t4new<-tnew^4

har.=harmonic(logBeer,5)
modelHR=lm(logBeer~har.+t+t2+t3+t4)
summary(modelHR)
plot(ts(fitted(modelHR),freq=12,start=c(1956,1)),ylab='log Beer',type='l',ylim=range(c(fitted(modelHR),logBeer))); 
points(logBeer)
residFromHR<-ts(residuals(modelHR), frequency=12, start=c(1956,1))
plot(residFromHR)

chosenMod<-modelHR
library(TSA)

har.=harmonic(log(beer_forecast),5)
newRegData<-data.frame(t=tnew, t2=t2new, t3=t3new, t4=t4new)
newRegData<-newRegData[(nrow(newRegData)-numToFor+1):nrow(newRegData),]
newRegData<-data.frame(har., newRegData)
colnames(newRegData)<-c(colnames(har.),"t","t2","t3","t4")
```

#Diagnostics on Cosine Trend
```{r}
adf.test(residFromHR)
pp.test(residFromHR)
acf(residFromHR)
pacf(residFromHR)
qqnorm(residFromHR)
qqline(residFromHR)
```


#Forecast based on cosine trend model
```{r}
modelString<-"Cosine trend model"
predictions<-predict(chosenMod,newdata =newRegData,se.fit = T)
pred<-ts(predictions$fit,start = c(1993,9),frequency = 12)
uci<-ts(pred+2*predictions$se.fit,start = c(1993,9),frequency = 12)
lci<-ts(pred-2*predictions$se.fit,start = c(1993,9),frequency = 12)
ymin=min(c(as.vector(lci),logBeer))-1
ymax=max(c(as.vector(uci),logBeer))+1
plot(logBeer,ylim=c(ymin,ymax),main=modelString, ylab='Logged Beer Production (Megalitres)')
lines(pred,col=2)
lines(uci,col=3)
lines(lci,col=3)

ymin=min(c(as.vector(lci),logBeer))-1
ymax=max(c(as.vector(uci),logBeer))+1
plot(logBeer,xlim=c(1993, 1996), ylim=c(4.5,5.5),main=modelString, ylab='Logged Beer Production (Megalitres)')
lines(pred,col=2)
lines(uci,col=3)
lines(lci,col=3)
lines(log(beer_forecast), col="black")

```


  
#Seasonal means model with ARIMA

```{r}
logBeer<-log(beerTS)
t<-1:length(logBeer)
t2<-t^2
t3<-t^3
t4<-t^4
month<-season(logBeer)

seasMod<-lm(beerTS~month)
summary(seasMod)

seas_resid<-ts(residuals(seasMod),frequency=12, start=c(1956,1)) 
plot(seas_resid, main="Residuals from Seasonal Means Model")
abline(h=0)


order4polyFit<-lm(beerTS~t+t2+t3+t4)
summary(order4polyFit)

#### plot the data and the fitted 4th order polynomial trend function
plot(x=1:length(beerTS),y=beerTS,type='l',ylab="Beer Production (Megalitres)",xlab="Time - Number of Months Since Jan 1956",main="order4poly Fit on Beer Production Data")
curve(expr = coef(order4polyFit)[1]+coef(order4polyFit)[2]*x+coef(order4polyFit)[3]*x^2+coef(order4polyFit)[4]*x^3+coef(order4polyFit)[5]*x^4,lty=1,add = TRUE, col="red")

poly_resid<-ts(residuals(order4polyFit),frequency=12, start=c(1956,1)) 
plot(poly_resid, main="Residuals from 4th order polynomial Model")
abline(h=0)

logSeasPoly<-lm(logBeer~t+t2+t3+t4+month)
summary(logSeasPoly)

logSeasPoly_resid<-ts(residuals(logSeasPoly),frequency=12, start=c(1956,1)) 
plot(logSeasPoly_resid, main="Residuals from Logged Beer\nSeasonal Means and 4th Order Poly Fit")
abline(h=0)
plot(logSeasPoly_resid, main="Residuals from Logged Beer\nSeasonal Means and 4th Order Poly Fit", type="l")
points(y=logSeasPoly_resid, x=time(logSeasPoly_resid), pch=as.vector(season(logSeasPoly_resid)))
abline(h=0)


acf(logSeasPoly_resid, main="ACF of Residuals from Logged Beer- Seas Means and Poly Fit", lag=35)
pacf(logSeasPoly_resid,main="PACF of Residuals from Logged Beer- Seas Means and Poly Fit", lag=35)

adf.test(logSeasPoly_resid)
pp.test(logSeasPoly_resid)

```

#Try an AR 12 model with the seasonal means and polynomial
```{r fig.width=8, fig.height=8}
#Set up external regressors and dummy vars
library(forecast)
monthDummies<-seasonaldummy(logBeer)
externReg<-data.frame(t, t2, t3, t4, monthDummies)

ar12_poly<-arima(logBeer, order=c(12,0,0), xreg=externReg)
ar12_poly
ar12_poly_resid<-ts(resid(ar12_poly), start=c(1956, 1), frequency=12)
plot(ar12_poly_resid, xlab="residual", ylab="time",main="Residuals from ARIMA(12,0,0) with Seas/Poly Deterministic Trend")
abline(h=0)
acf(ar12_poly_resid,main="ACF of Residuals from ARIMA(12,0,0)\nwith Seas/Poly Deterministic Trend")
pacf(ar12_poly_resid,main="PACF of Residuals from ARIMA(12,0,0)\nwith Seas/Poly Deterministic Trend",mar=c(4,2,2,0.8), oma=c(2,1,1,1))

```


#Build models based on the seasonal means only
```{r}
seasMeansMod<-lm(logBeer~season(logBeer))
plot(ts(residuals(seasMeansMod), start=c(1956,1),frequency=12), ylab="Residuals", main="Residual from Seasonal Means Fit")
abline(h=0)


ar12<-arima(logBeer, order=c(12,0,0), xreg=monthDummies)
ar12
tsdiag(ar12, gof.lag=20)

#residuals
ar12_resid<-ts(residuals(ar12), frequency=12, start=c(1956,1)) 
plot(ar12_resid, main="AR 12 model Residuals from Logged Beer\nSeasonal Means")
abline(h=0)
plot(ar12_resid, main="AR 12 model Residuals from Logged Beer\nSeasonal Means", type="l")
points(y=ar12_resid, x=time(ar12_resid), pch=as.vector(season(ar12_resid)))
abline(h=0)

acf(ar12_resid, main="ACF Plot of Residuals from ARIMA(12,0,0)")
pacf(ar12_resid, main="PACF Plot of Residuals from ARIMA(12,0,0)")

ar13<-arima(logBeer, order=c(13,0,0), xreg=monthDummies)
ar13
tsdiag(ar13, gof.lag=20)

ar12ma1<-arima(logBeer, order=c(12,0,1), xreg=monthDummies)
ar12ma1
tsdiag(ar12ma1, gof.lag=20)

ar12ma1_resid<-ts(residuals(ar12ma1), frequency=12, start=c(1956,1)) 
plot(ar12ma1_resid, main="AR 12 MA 1 model Residuals from Logged Beer\nseasonal Means", type="l")
points(y=ar12ma1_resid, x=time(ar12ma1_resid), pch=as.vector(season(ar12ma1_resid)))
abline(h=0)

acf(ar12ma1_resid,main="ACF of Residuals from ARIMA(12,0,1)\nwith Seas Trend")
pacf(ar12ma1_resid,main="PACF of Residuals from ARIMA(12,0,1)\nwith Seas Trend",mar=c(5,2,2,0.8), oma=c(1,1,1,1))

ar13ma1<-arima(logBeer, order=c(13,0,1), xreg=monthDummies)
ar13ma1
tsdiag(ar13ma1, gof.lag=20)

ar12ma2<-arima(logBeer, order=c(12,0,2), xreg=monthDummies)
ar12ma2
tsdiag(ar12ma2, gof.lag=20)

par(mfrow=c(1,1))

qqnorm(residuals(ar12), main="Normal QQ Plot of Residuals from ARIMA(12,0,0)")
qqline(ar12_resid)
shapiro.test(ar12_resid)


LB.test(ar12, lag=35)


pacf_acf<-data.frame(acfVal=acf(ar12_resid, plot=FALSE)$acf, pacfVal=pacf(ar12_resid, plot=FALSE)$acf)
#print(pacf_acf)

```


#Make the forecasts- set up external regressor data frame
```{r}
newMonthDummy<-seasonaldummy(beer_forecast)
```


#Plot the model forecasts
```{r}
chosenMod<-ar12
predictions<-predict(chosenMod, n.ahead=24,newxreg=newMonthDummy)
pred<-predictions$pred
uci<-pred+2*predictions$se
lci<-pred-2*predictions$se

sumSqErrSeasMean<-sum((log(beer_forecast)-pred)^2)

ymin=min(c(as.vector(lci),logBeer))-1
ymax=max(c(as.vector(uci),logBeer))+1
plot(logBeer,ylim=c(ymin,ymax),main=modelString, ylab='Logged Beer Production (Megalitres)')
lines(pred,col=2)
lines(uci,col=3)
lines(lci,col=3)

ymin=min(c(as.vector(lci),logBeer))-1
ymax=max(c(as.vector(uci),logBeer))+1
plot(logBeer,xlim=c(1993, 1996), ylim=c(4.5,5.5),main=modelString, ylab='Logged Beer Production (Megalitres)')
lines(pred,col=2)
lines(uci,col=3)
lines(lci,col=3)
lines(log(beer_forecast), col="black")

```





#SARIMA model with population data
```{r}
logBeer<-log(beerPopMonth$beer)
lag6_10_14<-beerPopMonth$lag6_10_14
t<-1:length(beerTS)
t2<-t^2
t3<-t^3
t4<-t^4
externalReg<-data.frame(lag6_10_14, t, t2, t3, t4)
monthlyPopModel_log<-lm(logBeer ~ lag6_10_14+t+t2+t3+t4)
monthlyPopOnly_log<-lm(logBeer ~ lag6_10_14)
summary(monthlyPopModel_log)

residFromPop_log<-ts(residuals(monthlyPopModel_log), frequency=12, start=c(1956,1))
plot(residFromPop_log, main="Plot of residuals from Pop/Poly Trend logged model")
abline(h=0)
plot(residFromPop_log, type="l", main="Plot of residuals from Pop/Poly Trend logged model")
points(y=residFromPop_log, x=time(residFromPop_log), pch=as.vector(season(residFromPop_log)))
abline(h=0)
```



```{r}
adf.test(residFromPop_log)
pp.test(residFromPop_log)
nsdiffs(residFromPop_log, m=frequency(residFromPop_log), test=c("ocsb","ch"), max.D=1)
acf(residFromPop_log)
pacf(residFromPop_log)
eacf(residFromPop_log)
```


#Build initial model
```{r}
popModel1<-arima(residFromPop_log, order=c(1,0,0), seasonal=list(order=c(1,0,0), period=12))
plotResid(popModel1)
```

#Try p=12 and q=12
```{r}
popModel2<-arima(residFromPop_log, order=c(12,0,0), seasonal=list(order=c(1,0,0), period=12))
plotResid(popModel2)
tsdiag(popModel2)

popModel3<-arima(residFromPop_log, order=c(1,0,12), seasonal=list(order=c(1,0,0), period=12))
plotResid(popModel3)
tsdiag(popModel3)
```


#Try overfitting the SARIMA(12,0,0)(1,0,0)[12]
```{r}
#popModel4<-arima(residFromPop_log, order=c(13,0,0), seasonal=list(order=c(1,0,0), period=12))
#Produces error

popModel5<-arima(residFromPop_log, order=c(12,0,1), seasonal=list(order=c(1,0,0), period=12))
popModel5

popModel6<-arima(residFromPop_log, order=c(12,0,0), seasonal=list(order=c(2,0,0), period=12))
popModel6

popModel7<-arima(residFromPop_log, order=c(12,0,0), seasonal=list(order=c(1,0,1), period=12))
popModel7
```

#Overfit model 7 (`r getModelString(popModel7)`)
```{r}
popModel8<-arima(residFromPop_log, order=c(13,0,0), seasonal=list(order=c(1,0,1), period=12))
popModel8

popModel9<-arima(residFromPop_log, order=c(12,0,1), seasonal=list(order=c(1,0,1), period=12))
popModel9

popModel10<-arima(residFromPop_log, order=c(12,0,0), seasonal=list(order=c(2,0,1), period=12))
popModel10

popModel11<-arima(residFromPop_log, order=c(12,0,0), seasonal=list(order=c(1,0,2), period=12))
popModel11
```

#Run diagnostics on model 7
```{r fig.width=7, fig.height=4}
chosenMod<-popModel7
modelString<-paste(getModelString(chosenMod), "with Population Data")

par(mfrow=c(1,1))
plot(residuals(chosenMod), main=paste("Residuals of Model", modelString), ylab="Residual")
abline(h=0)
par(mfrow=c(1,1))
plot(residuals(chosenMod), main=paste("Residuals of Model", modelString), ylab="Residual", type="l")
points(y=residuals(chosenMod), x=time(residuals(chosenMod)), pch=as.vector(season(residuals(chosenMod))))
abline(h=0)
```


```{r fig.width=7, fig.height=4}
par(mfrow=c(1,1))
qqnorm(residuals(chosenMod), main=paste("Normal QQ Plot of Residuals from", modelString))
qqline(residuals(chosenMod))

shapiro.test(residuals(chosenMod))

```


```{r fig.width=7, fig.height=4}
par(mfrow=c(1,1))
acf(residuals(chosenMod), main=paste("ACF of Residuals from", modelString))

LB.test(chosenMod, lag=35)
runs(residuals(chosenMod))
```


#Set up for forecast of population model
```{r}
startDate<-round(start(beer_forecast)[1]+start(beer_forecast)[2]/12,2)
endDate<-round(end(beer_forecast)[1]+end(beer_forecast)[2]/12,2)
numToFor<-length(beer_forecast)

allBeerData<-ts(c(beerTS, beer_forecast), start=c(1956, 1), frequency=12)

#Pull data for population
lag6_10_14new<-subset(monthlyLag6Long, as.numeric(rownames(monthlyLag6Long))>=startDate)
lag6_10_14new<-subset(lag6_10_14new, as.numeric(rownames(lag6_10_14new))<=endDate)

tnew<-1:length(allBeerData)
t2new<-tnew^2
t3new<-tnew^3
t4new<-tnew^4

newRegData<-data.frame(t=tnew, t2=t2new, t3=t3new, t4=t4new)
newRegData<-newRegData[(nrow(newRegData)-numToFor+1):nrow(newRegData),]
newRegData<-data.frame(lag6_10_14=lag6_10_14new, newRegData)
colnames(newRegData)<-c("lag6_10_14", colnames(newRegData)[2:5])

predFromPop<-predict(monthlyPopModel_log, newdata=newRegData, se.fit=TRUE)
uci_lm<-ts(predFromPop$fit+2*predFromPop$se.fit, start=c(1993, 9), frequency = 12)
lci_lm<-ts(predFromPop$fit-2*predFromPop$se.fit, start=c(1993, 9), frequency = 12)

ymin=min(c(as.vector(lci),logBeer))-1
ymax=max(c(as.vector(uci),logBeer))+1
plot(logBeer,xlim=c(1993, 1996), ylim=c(4.5,5.5),main=modelString, ylab='Logged Beer Production (Megalitres)')
lines(ts(predFromPop$fit, start=c(1993, 9), frequency = 12),col=2)
lines(uci_lm,col=3)
lines(lci_lm,col=3)
lines(log(beer_forecast), col="black")

```


#Plot the model forecasts

```{r}
 
predictions<-predict(chosenMod, n.ahead=24)

pred_raw<-(predictions$pred+predFromPop$fit)
pred<-predictions$pred+predFromPop$fit
uci<-pred_raw+2*predictions$se
lci<-pred_raw-2*predictions$se
uci_lm<-pred_raw+2*(predictions$se+predFromPop$se.fit)
lci_lm<-pred_raw-2*(predictions$se+predFromPop$se.fit)

sumSqErrSARIMA<-sum((log(beer_forecast)-pred)^2)

ymin=min(c(as.vector(lci),logBeer))-1
ymax=max(c(as.vector(uci),logBeer))+1
plot(logBeer,ylim=c(ymin,ymax),main=modelString, ylab='Logged Beer Production (Megalitres)')
lines(pred,col=2)
lines(uci,col=3)
lines(lci,col=3)
lines(log(beer_forecast), col="black")

ymin=min(c(as.vector(lci),log(beer_forecast)))-0.1
ymax=max(c(as.vector(uci),log(beer_forecast)))+0.1
plot(logBeer,xlim=c(1993, 1996), ylim=c(ymin,ymax),main=modelString, ylab='Logged Beer Production (Megalitres)')
lines(pred,col=2)
lines(uci,col=3)
lines(lci,col=3)
lines(uci_lm,col="blue")
lines(lci_lm,col="blue")
lines(log(beer_forecast), col="black")

```











